---
pagetitle: "Analysis_2025_01TFMLBlanc"
title: <!--html_preserve--><table width='100%'><tr><td><!--/html_preserve-->CONFIDENTIAL-NOT MENTION- NOT SPREAD<!--html_preserve--></td><td align='right'><!--/html_preserve-->2025_01TFMLBlanc- Small Scenario Versió 4 <!--html_preserve--></td></tr></table> <!--/html_preserve-->
subtitle: <!--html_preserve--><h6><!--/html_preserve-->**Statistical report TFM Lucia Blanc**<!--html_preserve--></h6><!--/html_preserve-->
author:
- name: "**Research team - ** Lucia Blanc"
- name: "UOC"
- name: "**Product - ** Statistical report"
- name: "**Biostatistics Research Unit HUGTiP-IGTP - ** Lucia Blanc"

date: today
execute:
  echo: false
  warning: false
  message: false
format: 
  html:
    output-file: 'Analysis_2025_01TFMLBlanc_small_v4_loop_v2.html'
    embed-resources: true
    toc: true
    toc_depth: 5
    toc_float:
      collapsed: true
      smooth_scroll: true
    ##Estil 
    theme: cosmo 
    highlight: tango
    css: markdown.css
    ###número seccions
    number-sections: true
params:
  report_seed: 05052025
---

```{r}
#| results: 'hide'
#| include: FALSE

# rm(list=ls())

# #Package renv (to restore package versions):
# codiR <- 'I:/CTebe/2_Projectes/2025_01TFMLBlanc/3_CodiR'
# renv::init(codiR) #To generate a copy of the package versions (first time)
# renv::snapshot(codiR) #To regenerate the copy if we change or update any library (at the end of the analysis)
# renv::restore(codiR) #To restore the package versions (for the future, we would have to create a project for this)

#Load libraries
library(kableExtra) 
library(dplyr) 
library(tidyr) 
library(tibble)
library(purrr)
library(stringr)
library(epiR)
library(captioner) 
library(lme4) 
library(sjPlot)
library(ggplot2)
library(gtsummary)
library(gt)
library(flowchart)
library(cutpointr)
library(pROC)
library(cutpointr)
library(DescTools)
library(OptimalCutpoints)
library(ThresholdROC)
###############

#Dades
################
setwd('I:/CTebe/2_Projectes/2025_01TFMLBlanc')
dades_ana<-'./2_Dades/2_Analisi'
Funcions<-'I:/CTebe/3_CodiR/Funcions'
Themes<-'I:/CTebe/3_CodiR/Themes'

# Enumerate tables and figures
source(file.path(Funcions, 'captions_en.R'))

#Source themes:
source(file.path(Themes, 'gtsummary_theme_bru_v2.R'))

#Set the theme
reset_gtsummary_theme()
set_gtsummary_theme(theme_bru)
#Read data
# load(file.path(dades_ana, 'data_v1.Rda'))

#Get labels (to restore them if they are lost)
# labels <- labelled::get_variable_labels(data)
# data <- data %>%
#    labelled::set_variable_labels(.labels = labels, .strict = FALSE) #restore labels

#Read label dictionary
#dictionary <- readxl::read_excel(file.path(dades_ana, 'dictionary.xlsx'))

nboot<- 50
```


```{r}
print(params$report_seed)
```

# Objectives

# Methodology

Study variables will be described in tables by study group. Categorical variables will be described by the frequency and percentage of each category. Continuous variables will be described by the mean and standard deviation or by the median and interquantile range, depending on the distribution of the variables.

The conditions of use of the models were validated and, whenever possible, confidence intervals at 95% were calculated. All analyses were performed with the statistical program R version 4.4.0 (2024-06-14) for Windows.


# Simulation data

# Normal distribution: Small scenario
```{r echo=FALSE, message=FALSE, warning=FALSE}
# set.seed(123)
set.seed(params$report_seed)
# Paràmetres

n <- 1000 # observacions
mu <- 0.5  # constant
sigma <- 1  # sd de l'error
prev <- 0.1

x <- rbinom(n, 1, 0.1)


dades_ep <- tibble(x = x) %>%
  group_by(x) |> 
  mutate(
    y = case_when(
      x == 1 ~ rnorm(n(), mean = mu, sd = sigma),
      x == 0 ~ rnorm(n(), mean = 0, sd = 1)
    )
  )


# Guardar dataframes

# dades <- list()
# save(dades, file = "dades_loop_small_v2.Rdata")

load("dades_loop_small_v2.Rdata")

var_seed <- params$report_seed

dades <- dades |> purrr::list_assign(var_seed = dades_ep)

names(dades)[names(dades) == "var_seed"] <- var_seed

save(dades, file = "dades_loop_small_v2.Rdata")

# Dades simulades
# head(dades_ep)

# tapply(dades_ep$y, dades_ep$x, mean)
# tapply(dades_ep$y, dades_ep$x, sd)

```


&nbsp; 

`r tab_cap("Description of data simulation: Small scenario")`

```{r echo=FALSE, message=FALSE, warning=FALSE}

dades_ep |> 
  dplyr::select(x, y) |> 
  gtsummary::tbl_summary(
    type = x ~"categorical",
        digits = all_continuous() ~ 2)


model <- lm(y ~ x,data=dades_ep)

tab_model(model,show.se=T)
```

&nbsp; 
&nbsp; 

`r fig_cap("Density comparison plot")`
```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}

dades_ep$x_c<-factor(dades_ep$x,
                    levels = c(0,1),
                    labels = c("Non-disease", "Disease"))

dades_ep$y_pred<-dades_ep$y

d <- dplyr::select(dades_ep,c(y_pred, x_c, x))
d <- filter(d, x_c%in%c("Non-disease", "Disease"))
d$x_c<-factor(d$x_c, levels = c("Non-disease", "Disease"))


#png(r"(I:\CTebe\2_Projectes\2025_01TFMLBlanc\4_Productes\Presentation\WWW\small_density.png)",width = 5000,height = 4000,res = 600)

ggplot(d, aes(y_pred,fill=x_c,color=x_c)) +
 geom_density(alpha = 0.1) +
   geom_vline(xintercept = mean(d$y_pred), linetype = "dashed", color = "black", linewidth = 1.2) +
 geom_vline(xintercept = mean(d$y_pred[d$x_c=="Non-disease"]), linetype = "dashed", color = "red", linewidth = 1.2,alpha = 0.8) +
 geom_vline(xintercept = mean(d$y_pred[d$x_c=="Disease"]), linetype = "dashed", color = "blue", linewidth = 1.2,alpha = 0.8) +
 theme_minimal()+labs(x="Predictor")+
  ggtitle("Density comparison for null scenario")+ 
  guides(fill = guide_legend(title = "Binary outcome"), 
        legend.background = element_rect(fill = "white"), colour = guide_legend(title = "Binary outcome"))
      
```

&nbsp; 
&nbsp; 

`r fig_cap("ROC Curve")`
```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}

#png(r"(I:\CTebe\2_Projectes\2025_01TFMLBlanc\4_Productes\Presentation\WWW\small_roc.png)",width = 5000,height = 4000,res = 600)


myroc <- roc(d$x, d$y_pred, ci=TRUE)
plot(myroc, main="ROC curve and AUC for small scenario", xlim = c(1, 0), lwd=3)
text(0.4, 0.1, paste0("AUC:", round(auc(myroc), 2), 
                      " [",round(myroc$ci[1],2), ", ",round(myroc$ci[2],2),"]" ))

```

&nbsp; 
&nbsp; 

## Methods

### New approach

```{r echo=FALSE, message=FALSE, warning=FALSE}
d_cut <- data.frame(
  x = as.numeric(d$x),
  y_pred = as.numeric(d$y_pred)
)
```

#### Bootstrapping

```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}

# Bootstrapping
 # set.seed(123)
 boots<-seq(1:nrow(d))
 p<-c()
 for(j in 1:nboot){
   # print(j)
   llista_s <- sample(boots, replace=T)
   ds <- d[llista_s,]
   x_outcome <-quantile(d$y_pred,probs = seq(0,1,0.01),na.rm=T)

   ai <- c()
   bi <- c()
   val <- c()
   hr <- c()
   p1x <- c()
   p2x <- c()
   cind <- c()
   lr <- c()
   br <- c()
   pvalue <- c()
   voltes <- c()
   for(i in x_outcome){
     ds$y <- cut(ds$y_pred, breaks = c(-Inf,i, Inf))
     if(all(table(ds$y, ds$x)>1)){
       m <- glm(x~y,data=ds, family=poisson(link = "log"))
       l <- summary(m)
       pvalue <- c(pvalue, l$coefficients[2,4])
       ai <- c(ai,AIC(m))
       bi <- c(bi, BIC(m))
       val <- c(val,i)
       h <- exp(l$coefficients[2,1])
       p1 <- round(exp(confint(m))[2,1],2)
       p2 <- round(exp(confint(m))[2,1],2)
       hr <- c(hr,h)
       p1x <- c(p1x, p1)
       p2x <- c(p2x, p2)
       br <- c(br, BrierScore(m))
       voltes <- c(voltes, j)
     }
   }
   d3 <- data.frame(voltes, val, hr, p1x,p2x,pvalue,br)

   minim <- min(d3$pvalue[d3$voltes==j])
   coff <- min(d3[d3$pvalue==minim& d3$voltes==j, "val"])
    print(coff)
    p<-rbind(p,coff)
 }
 
  p<-as.data.frame(p)
 
# save(p, file = file.path("I:\CTebe\2_Projectes\2025_01TFMLBlanc\2_Dades\2_Analisi", "Null_scenario.Rda"))

 
# load(file="I:/CTebe/2_Projectes/2025_01TFMLBlanc/2_Dades/2_Analisi/Small/Small_scenario.Rda")

```

&nbsp; 

`r fig_cap("Density function with high-density local maxima")`

```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
funcio_densitat <- density(p$V1)
pics <- which(diff(sign(diff(funcio_densitat$y))) == -2) + 1
maximlocal_x <- funcio_densitat$x[pics]
maximlocal_y <- funcio_densitat$y[pics]

data.frame(valors=round(maximlocal_x,2),densitat=round(maximlocal_y,2))%>%
  gt()
  

ggplot(data.frame(x = funcio_densitat$x, y = funcio_densitat$y), aes(x = x, y = y)) +
  geom_line(color = "blue") +  
  geom_point(data = data.frame(x = maximlocal_x[maximlocal_y>mean(funcio_densitat$y)],
                               y = maximlocal_y[maximlocal_y>mean(funcio_densitat$y)]),
             aes(x = x, y = y), col = "red", pch = 16)+  
  geom_vline(aes(xintercept=mean(p$V1[p$V1!="Inf"]), color="orange"),show_guide=TRUE)+
  labs(title = "Density function with high-density local maxima",
       x = "X-axis",y = "Density")+
  scale_color_manual("Cut-off", values=c(orange="orange"),labels=c( "E(values)"))+
  theme_bw()
```
&nbsp; 
&nbsp; 

##### Expected value 𝐸[𝑋]
```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}

xintercept=mean(p$V1[p$V1!="Inf"])

#construim dicotòmica
d$y_cat_b <- ifelse(d$y_pred>xintercept, "Positive", "Negative")

#construim la taula
tab_b <- table(d$y_cat_b, d$x)
tab_b <- tab_b[c(2,1),c(2,1)]
x <- diagnostic(tab_b)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
y <- epiR::epi.tests(tab_b)
y
```
&nbsp; 

The diagnostic indices taking `r round(xintercept,3)` as the cut-off point are as follows 

```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
kable(round(x[c(1:4,9),],3), caption = "Predictor", align = "l") %>%
 kable_styling(bootstrap_options = c("striped","hover"))

### Matriz de confusión (paquete caret) - https://rdrr.io/cran/caret/man/confusionMatrix.html
d$y_catfn_b <- factor(d$y_cat_b,
                   levels=c("Positive", "Negative"),
                   labels=c("Non-disease", "Disease"))
machine_matriz <- caret::confusionMatrix(d$y_catfn_b, d$x_c)


metrics_boot_expect <- data.frame(
  method = "New approach: Expected value",
  optimal_cutpoint = round(xintercept, 3),
  sensitivity = x["Sensitivity", "Estim."],
  sensitivity_low = x["Sensitivity", "Low.lim(95%)"],
  sensitivity_high = x["Sensitivity", "Up.lim(95%)"],
  
  specificity = x["Specificity", "Estim."],
  specificity_low = x["Specificity", "Low.lim(95%)"],
  specificity_high = x["Specificity", "Up.lim(95%)"],
  
  ppv = x["Pos.Pred.Val.", "Estim."],
  ppv_low = x["Pos.Pred.Val.", "Low.lim(95%)"],
  ppv_high = x["Pos.Pred.Val.", "Up.lim(95%)"],
  
  npv = x["Neg.Pred.Val.", "Estim."],
  npv_low = x["Neg.Pred.Val.", "Low.lim(95%)"],
  npv_high = x["Neg.Pred.Val.", "Up.lim(95%)"],
  
  accuracy = x["Accuracy", "Estim."],
  accuracy_low = x["Accuracy", "Low.lim(95%)"],
  accuracy_high = x["Accuracy", "Up.lim(95%)"]
)
```

&nbsp; 
&nbsp; 


##### Maximum density
```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}

# x con mayor densidad 
xintercept <- maximlocal_x[which.max(maximlocal_y)]


#construim dicotòmica
d$y_cat_b <- ifelse(d$y_pred>xintercept, "Positive", "Negative")

#construim la taula
tab_b <- table(d$y_cat_b, d$x)
tab_b <- tab_b[c(2,1),c(2,1)]
x <- diagnostic(tab_b)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
y <- epiR::epi.tests(tab_b)
y
```
&nbsp; 

The diagnostic indices taking `r round(xintercept,3)` as the cut-off point are as follows 

```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
kable(round(x[c(1:4,9),],3), caption = "Predictor", align = "l") %>%
 kable_styling(bootstrap_options = c("striped","hover"))

### Matriz de confusión (paquete caret) - https://rdrr.io/cran/caret/man/confusionMatrix.html
d$y_catfn_b <- factor(d$y_cat_b,
                   levels=c("Positive", "Negative"),
                   labels=c("Non-disease", "Disease"))
machine_matriz <- caret::confusionMatrix(d$y_catfn_b, d$x_c)


metrics_boot_maxdens <- data.frame(
  method = "New approach: Maximum density",
  optimal_cutpoint = round(xintercept, 3),
  sensitivity = x["Sensitivity", "Estim."],
  sensitivity_low = x["Sensitivity", "Low.lim(95%)"],
  sensitivity_high = x["Sensitivity", "Up.lim(95%)"],
  
  specificity = x["Specificity", "Estim."],
  specificity_low = x["Specificity", "Low.lim(95%)"],
  specificity_high = x["Specificity", "Up.lim(95%)"],
  
  ppv = x["Pos.Pred.Val.", "Estim."],
  ppv_low = x["Pos.Pred.Val.", "Low.lim(95%)"],
  ppv_high = x["Pos.Pred.Val.", "Up.lim(95%)"],
  
  npv = x["Neg.Pred.Val.", "Estim."],
  npv_low = x["Neg.Pred.Val.", "Low.lim(95%)"],
  npv_high = x["Neg.Pred.Val.", "Up.lim(95%)"],
  
  accuracy = x["Accuracy", "Estim."],
  accuracy_low = x["Accuracy", "Low.lim(95%)"],
  accuracy_high = x["Accuracy", "Up.lim(95%)"]
)
```

&nbsp; 
&nbsp; 


### Youden's J statistic

```{r echo=FALSE, message=FALSE, warning=FALSE}

youden_res <- optimal.cutpoints(
  X = "y_pred",
  status = "x", 
  tag.healthy = 0,
  methods = "Youden",
  data = d_cut, 
  ci.fit = TRUE,
  conf.level = 0.95,
  trace = FALSE
)


summary(youden_res)

# Metrics results
opt_youden <- summary(youden_res)$Youden$Global$optimal.cutoff

cutoff_youden <- opt_youden$cutoff

se_youden<-opt_youden$Se
sp_youden<-opt_youden$Sp
ppv_youden<-opt_youden$PPV
npv_youden<-opt_youden$NPV

# Càlculs CI
sensitivity <- se_youden[1, "Value"]
sensitivity_low <-se_youden[1, "ll"]
sensitivity_high <- se_youden[1, "ul"]

specificity <- sp_youden[1, "Value"]
specificity_low <- sp_youden[1, "ll"]
specificity_high <- sp_youden[1, "ul"]
  
ppv <- ppv_youden[1, "Value"]
ppv_low <- ppv_youden[1, "ll"]
ppv_high <- ppv_youden[1, "ul"]
  
npv <- npv_youden[1, "Value"]
npv_low <- npv_youden[1, "ll"]
npv_high <- npv_youden[1, "ul"]



# Classification

# Confusion matrix
d$y_cat <- ifelse(d$y_pred >= cutoff_youden, "Predicted disease", "Predicted non-disease")
tab <- table(d$y_cat, d$x_c)
tab <- tab[,c(2,1)]

kable(tab, caption = "Contingency Table - Youden", align = "c") %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                 full_width = F, position = "center")

# Diagnostic
diag <- epi.tests(tab)
diag


# Accuracy
correct<-sum(tab[1,1], tab[2,2])

total <- sum(tab)

acc_youden <- as.numeric(correct/total)
ci_acc <- binom.test(correct, total, conf.level = 0.95)$conf.int

# Metrics
metrics_youden <- data.frame(
  method = "Youden",
  optimal_cutpoint = cutoff_youden,
  
  sensitivity = sensitivity,
  sensitivity_low = sensitivity_low,
  sensitivity_high = sensitivity_high,

  specificity = specificity,
  specificity_low = specificity_low,
  specificity_high = specificity_high,

  ppv = ppv,
  ppv_low = ppv_low,
  ppv_high = ppv_high,

  npv = npv,
  npv_low = npv_low,
  npv_high = npv_high,

  accuracy = acc_youden,
  accuracy_low = ci_acc[1],
  accuracy_high = ci_acc[2]
)


metrics_t<-metrics_youden %>%
   select(method, optimal_cutpoint, sensitivity, sensitivity_low, sensitivity_high,
          specificity, specificity_low, specificity_high, ppv, ppv_low, ppv_high, npv, npv_low, npv_high, accuracy, accuracy_low, accuracy_high) %>%
   gt() %>%
   tab_header(title = "Optimal Cut-off Point Metrics - Youden") %>%
   fmt_number(
     columns = where(is.numeric),
     decimals = 3
   )
```

&nbsp; 
&nbsp; 

### Euclidean distance
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Se and Sp
Se <- myroc$sensitivities
Sp <- myroc$specificities
thres <- myroc$thresholds

eucl_res <- sqrt((1 - Se)^2 + (1 - Sp)^2)
best_index <- which.min(eucl_res)

optimal_threshold <- thres[best_index]


# Classification

# Confusion matrix
d$y_cat <- ifelse(d$y_pred >=optimal_threshold,"Predicted disease", "Predicted non-disease")

tab <- table(d$y_cat, d$x_c)
tab <- tab[,c(2,1)]

kable(tab, caption = "Contingency Table - Youden", align = "c") %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                 full_width = F, position = "center")

# Diagnostic
diag <- epi.tests(tab)
diag

metrics<-diag[["detail"]]


# Càlcul CI
# PPV, NPV, Acc
se <- metrics[metrics$statistic == "se", ]
sp <- metrics[metrics$statistic == "sp", ]
ppv  <- metrics[metrics$statistic == "pv.pos", ]
npv  <- metrics[metrics$statistic == "pv.neg", ]
acc  <- metrics[metrics$statistic == "diag.ac", ]


# Metrics
metrics_eucl <- data.frame(
  method = "Euclidean distance",
  optimal_cutpoint = optimal_threshold,
  
  sensitivity = se$est,
  sensitivity_low = se$lower,
  sensitivity_high = se$upper,
  
  specificity = sp$est,
  specificity_low = sp$lower,
  specificity_high = sp$upper,
  
  ppv = ppv$est,
  ppv_low = ppv$lower,
  ppv_high = ppv$upper,
  
  npv = npv$est,
  npv_low = npv$lower,
  npv_high = npv$upper,
  
  accuracy = acc$est,
  accuracy_low = acc$lower,
  accuracy_high = acc$upper
)



metrics_t<-metrics_eucl %>%
   select(method, optimal_cutpoint, sensitivity, sensitivity_low, sensitivity_high,
          specificity, specificity_low, specificity_high, ppv, ppv_low, ppv_high, npv, npv_low, npv_high, accuracy, accuracy_low, accuracy_high) %>%
   gt() %>%
   tab_header(title = "Optimal Cut-off Point Metrics - Euclidean distance") %>%
   fmt_number(
     columns = where(is.numeric),
     decimals = 3
   )


```

&nbsp; 
&nbsp; 

### Maximum product of Se and Sp
```{r echo=FALSE, message=FALSE, warning=FALSE}

mprod_res <- optimal.cutpoints(
  X = "y_pred",
  status = "x", 
  tag.healthy = 0,
  methods = "MaxProdSpSe",
  data = d_cut, 
  ci.fit = TRUE,
  conf.level = 0.95,
  trace = FALSE
)



summary(mprod_res)

# Results
opt_mprod <- summary(mprod_res)$MaxProdSpSe$Global$optimal.cutoff

cutoff_mprod <- opt_mprod$cutoff

se_mprod<-opt_mprod$Se
sp_mprod<-opt_mprod$Sp
ppv_mprod<-opt_mprod$PPV
npv_mprod<-opt_mprod$NPV

# Càlculs CI
sensitivity <- se_mprod[1, "Value"]
sensitivity_low <-se_mprod[1, "ll"]
sensitivity_high <- se_mprod[1, "ul"]

specificity <- sp_mprod[1, "Value"]
specificity_low <- sp_mprod[1, "ll"]
specificity_high <- sp_mprod[1, "ul"]
  
ppv <- ppv_mprod[1, "Value"]
ppv_low <- ppv_mprod[1, "ll"]
ppv_high <- ppv_mprod[1, "ul"]
  
npv <- npv_mprod[1, "Value"]
npv_low <- npv_mprod[1, "ll"]
npv_high <- npv_mprod[1, "ul"]



# Classification

# Confusion matrix
d$y_cat <- ifelse(d$y_pred >= cutoff_mprod, "Predicted disease", "Predicted non-disease")


tab <- table(d$y_cat, d$x_c)
tab <- tab[,c(2,1)]

kable(tab, caption = "Contingency Table - Maximum product Se*Sp", align = "c") %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                 full_width = F, position = "center")

# Diagnostic
diag <- epi.tests(tab)
diag


# Accuracy
correct<-sum(tab[1,1], tab[2,2])

total <- sum(tab)

acc_mprod <- as.numeric(correct/total)
ci_acc <- binom.test(correct, total, conf.level = 0.95)$conf.int

# Metrics
metrics_mprod <- data.frame(
  method = "Maximum product Se*Sp",
  optimal_cutpoint = cutoff_mprod,
  
  sensitivity = sensitivity,
  sensitivity_low = sensitivity_low,
  sensitivity_high = sensitivity_high,

  specificity = specificity,
  specificity_low = specificity_low,
  specificity_high = specificity_high,

  ppv = ppv,
  ppv_low = ppv_low,
  ppv_high = ppv_high,

  npv = npv,
  npv_low = npv_low,
  npv_high = npv_high,

  accuracy = acc_mprod,
  accuracy_low = ci_acc[1],
  accuracy_high = ci_acc[2]
)


metrics_t<-metrics_mprod %>%
   select(method, optimal_cutpoint, sensitivity, sensitivity_low, sensitivity_high,
          specificity, specificity_low, specificity_high, ppv, ppv_low, ppv_high, npv, npv_low, npv_high, accuracy, accuracy_low, accuracy_high) %>%
   gt() %>%
   tab_header(title = "Optimal Cut-off Point Metrics - Max. product Se*Sp") %>%
   fmt_number(
     columns = where(is.numeric),
     decimals = 3
   )

```
&nbsp; 
&nbsp; 

### Index of union (IU) 
```{r echo=FALSE, message=FALSE, warning=FALSE}

#Càlcul AUC
auc_val <- as.numeric(auc(myroc)) 


roc_df <- data.frame( 
  threshold = myroc$thresholds, 
  sens = myroc$sensitivities, 
  spec = myroc$specificities 

) 

#Càlcul IU
roc_df$IU <- abs(roc_df$sens - auc_val) + abs(roc_df$spec - auc_val) 
best_iu <- roc_df[which.min(roc_df$IU), ] 



# Classification

# Confusion matrix
d$y_cat <- ifelse(d$y_pred >= best_iu$threshold, "Predicted disease", "Predicted non-disease")


tab <- table(d$y_cat, d$x_c)
tab <- tab[,c(2,1)]

kable(tab, caption = "Contingency Table - Index of Union", align = "c") %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                 full_width = F, position = "center")

# Diagnostic
diag <- epi.tests(tab)
diag


x_iu<-diagnostic(tab)

# Càlcul mètriques amb CI
se <- x_iu["Sensitivity", ]
sp <- x_iu["Specificity", ]
ppv  <- x_iu["Pos.Pred.Val.", ]
npv  <- x_iu["Neg.Pred.Val.", ]
acc  <- x_iu["Accuracy", ]


# Metrics
metrics_iu <- data.frame(
  method = "Index of Union (IU)",
  optimal_cutpoint = best_iu$threshold,
  
  sensitivity = se[1],
  sensitivity_low = se[2],
  sensitivity_high = se[3],
  
  specificity = sp[1],
  specificity_low = sp[2],
  specificity_high = sp[3],
  
  ppv = ppv[1],
  ppv_low = ppv[2],
  ppv_high = ppv[3],
  
  npv = npv[1],
  npv_low = npv[2],
  npv_high = npv[3],
  
  accuracy = acc[1],
  accuracy_low = acc[2],
  accuracy_high = acc[3]
)


metrics_t<-metrics_iu %>%
   select(method, optimal_cutpoint, sensitivity, sensitivity_low, sensitivity_high,
          specificity, specificity_low, specificity_high, ppv, ppv_low, ppv_high, npv, npv_low, npv_high, accuracy, accuracy_low, accuracy_high) %>%
   gt() %>%
   tab_header(title = "Optimal Cut-off Point Metrics - Index of Union") %>%
   fmt_number(
     columns = where(is.numeric),
     decimals = 3
   )

```

&nbsp; 
&nbsp; 




### Cost approach
```{r echo=FALSE, message=FALSE, warning=FALSE}

cost_ap_res <- optimal.cutpoints(
  X = "y_pred",
  status = "x", 
  tag.healthy = 0,
  methods = "CB",
  data = d_cut, 
  control = control.cutpoints(
  costs.ratio = 5), # It costs 5 times more to predict a FN than a FP
  ci.fit = TRUE,
  conf.level = 0.95,
  trace = FALSE
)

summary(cost_ap_res)

# Results
opt_cb <- summary(cost_ap_res)$CB$Global$optimal.cutoff

cutoff_cb <- opt_cb$cutoff

se_cb<-opt_cb$Se
sp_cb<-opt_cb$Sp
ppv_cb<-opt_cb$PPV
npv_cb<-opt_cb$NPV

# Càlculs CI
sensitivity <- se_cb[1, "Value"]
sensitivity_low <-se_cb[1, "ll"]
sensitivity_high <- se_cb[1, "ul"]

specificity <- sp_cb[1, "Value"]
specificity_low <- sp_cb[1, "ll"]
specificity_high <- sp_cb[1, "ul"]
  
ppv <- ppv_cb[1, "Value"]
ppv_low <- ppv_cb[1, "ll"]
ppv_high <- ppv_cb[1, "ul"]
  
npv <- npv_cb[1, "Value"]
npv_low <- npv_cb[1, "ll"]
npv_high <- npv_cb[1, "ul"]



# Classification

# Confusion matrix
d$y_cat <- ifelse(d$y_pred >= cutoff_cb, "Predicted disease", "Predicted non-disease")


tab <- table(d$y_cat, d$x_c)
tab <- tab[,c(2,1)]

kable(tab, caption = "Contingency Table - Cost approach", align = "c") %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                 full_width = F, position = "center")

# Diagnostic
diag <- epi.tests(tab)
diag


# Accuracy
correct<-sum(tab[1,1], tab[2,2])

total <- sum(tab)

acc_cb <- as.numeric(correct/total)
ci_acc <- binom.test(correct, total, conf.level = 0.95)$conf.int

# Metrics
metrics_cost <- data.frame(
  method = "Cost approach",
  optimal_cutpoint = cutoff_cb,
  
  sensitivity = sensitivity,
  sensitivity_low = sensitivity_low,
  sensitivity_high = sensitivity_high,

  specificity = specificity,
  specificity_low = specificity_low,
  specificity_high = specificity_high,

  ppv = ppv,
  ppv_low = ppv_low,
  ppv_high = ppv_high,

  npv = npv,
  npv_low = npv_low,
  npv_high = npv_high,

  accuracy = acc_cb,
  accuracy_low = ci_acc[1],
  accuracy_high = ci_acc[2]
)

metrics_t<-metrics_cost %>%
   select(method, optimal_cutpoint, sensitivity, sensitivity_low, sensitivity_high,
          specificity, specificity_low, specificity_high, ppv, ppv_low, ppv_high, npv, npv_low, npv_high, accuracy, accuracy_low, accuracy_high) %>%
   gt() %>%
   tab_header(title = "Optimal Cut-off Point Metrics - Cost approach") %>%
   fmt_number(
     columns = where(is.numeric),
     decimals = 3
   )

```

&nbsp; 
&nbsp; 






### Missclassification cost

```{r echo=FALSE, message=FALSE, warning=FALSE}

misscost_res <- optimal.cutpoints(
  X = "y_pred",
  status = "x", 
  tag.healthy = 0,
  methods = "MCT",
  data = d_cut, 
  ci.fit = TRUE,
  conf.level = 0.95,
  trace = FALSE
)


summary(misscost_res)

# Results
opt_misscost <- summary(misscost_res)$MCT$Global$optimal.cutoff

cutoff_misscost <- opt_misscost$cutoff

se_misscost<-opt_misscost$Se
sp_misscost<-opt_misscost$Sp
ppv_misscost<-opt_misscost$PPV
npv_misscost<-opt_misscost$NPV

# Càlculs CI
sensitivity <- se_misscost[1, "Value"]
sensitivity_low <-se_misscost[1, "ll"]
sensitivity_high <- se_misscost[1, "ul"]

specificity <- sp_misscost[1, "Value"]
specificity_low <- sp_misscost[1, "ll"]
specificity_high <- sp_misscost[1, "ul"]
  
ppv <- ppv_misscost[1, "Value"]
ppv_low <- ppv_misscost[1, "ll"]
ppv_high <- ppv_misscost[1, "ul"]
  
npv <- npv_misscost[1, "Value"]
npv_low <- npv_misscost[1, "ll"]
npv_high <- npv_misscost[1, "ul"]



# Classification

# Confusion matrix
d$y_cat <- ifelse(d$y_pred >= cutoff_misscost, "Predicted disease", "Predicted non-disease")


tab <- table(d$y_cat, d$x_c)
tab <- tab[,c(2,1)]

kable(tab, caption = "Contingency Table - Missclassification cost", align = "c") %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                 full_width = F, position = "center")

# Diagnostic
diag <- epi.tests(tab)
diag


# Accuracy
correct<-sum(tab[1,1], tab[2,2])

total <- sum(tab)

acc_misscost <- as.numeric(correct/total)
ci_acc <- binom.test(correct, total, conf.level = 0.95)$conf.int

# Metrics
metrics_misscost <- data.frame(
  method = "Missclassification cost",
  optimal_cutpoint = cutoff_misscost,
  
  sensitivity = sensitivity,
  sensitivity_low = sensitivity_low,
  sensitivity_high = sensitivity_high,

  specificity = specificity,
  specificity_low = specificity_low,
  specificity_high = specificity_high,

  ppv = ppv,
  ppv_low = ppv_low,
  ppv_high = ppv_high,

  npv = npv,
  npv_low = npv_low,
  npv_high = npv_high,

  accuracy = acc_misscost,
  accuracy_low = ci_acc[1],
  accuracy_high = ci_acc[2]
)



metrics_t<-metrics_misscost %>%
   select(method, optimal_cutpoint, sensitivity, sensitivity_low, sensitivity_high,
          specificity, specificity_low, specificity_high, ppv, ppv_low, ppv_high, npv, npv_low, npv_high, accuracy, accuracy_low, accuracy_high) %>%
   gt() %>%
   tab_header(title = "Optimal Cut-off Point Metrics - Missclassification cost") %>%
   fmt_number(
     columns = where(is.numeric),
     decimals = 3
   )

```

&nbsp; 
&nbsp; 

### Diagnosis Odds Ratio (DOR)
```{r echo=FALSE, message=FALSE, warning=FALSE}

dor_res <- optimal.cutpoints(
  X = "y_pred",
  status = "x", 
  tag.healthy = 0,
  methods = "MaxDOR",
  data = d_cut, 
  ci.fit = TRUE,
  conf.level = 0.95,
  trace = FALSE
)

summary(dor_res)

# Results
opt_dor <- summary(dor_res)$MaxDOR$Global$optimal.cutoff

cutoff_dor <- opt_dor$cutoff

se_dor<-opt_dor$Se
sp_dor<-opt_dor$Sp
ppv_dor<-opt_dor$PPV
npv_dor<-opt_dor$NPV

# Càlculs CI
sensitivity <- se_dor[1, "Value"]
sensitivity_low <-se_dor[1, "ll"]
sensitivity_high <- se_dor[1, "ul"]

specificity <- sp_dor[1, "Value"]
specificity_low <- sp_dor[1, "ll"]
specificity_high <- sp_dor[1, "ul"]
  
ppv <- ppv_dor[1, "Value"]
ppv_low <- ppv_dor[1, "ll"]
ppv_high <- ppv_dor[1, "ul"]
  
npv <- npv_dor[1, "Value"]
npv_low <- npv_dor[1, "ll"]
npv_high <- npv_dor[1, "ul"]



# Classification

# Confusion matrix
d$y_cat <- ifelse(d$y_pred >= cutoff_dor, "Predicted disease", "Predicted non-disease")


tab <- table(d$y_cat, d$x_c)
tab <- tab[,c(2,1)]

kable(tab, caption = "Contingency Table - Diagnostic Odds Ratio", align = "c") %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                 full_width = F, position = "center")

# Diagnostic
diag <- epi.tests(tab)
diag


# Accuracy
correct<-sum(tab[1,1], tab[2,2])

total <- sum(tab)

acc_dor <- as.numeric(correct/total)
ci_acc <- binom.test(correct, total, conf.level = 0.95)$conf.int

# Metrics
metrics_dor <- data.frame(
  method = "Diagnostic Odds Ratio",
  optimal_cutpoint = cutoff_dor,
  
  sensitivity = sensitivity,
  sensitivity_low = sensitivity_low,
  sensitivity_high = sensitivity_high,

  specificity = specificity,
  specificity_low = specificity_low,
  specificity_high = specificity_high,

  ppv = ppv,
  ppv_low = ppv_low,
  ppv_high = ppv_high,

  npv = npv,
  npv_low = npv_low,
  npv_high = npv_high,

  accuracy = acc_dor,
  accuracy_low = ci_acc[1],
  accuracy_high = ci_acc[2]
)


metrics_t<-metrics_dor %>%
   select(method, optimal_cutpoint, sensitivity, sensitivity_low, sensitivity_high,
          specificity, specificity_low, specificity_high, ppv, ppv_low, ppv_high, npv, npv_low, npv_high, accuracy, accuracy_low, accuracy_high) %>%
   gt() %>%
   tab_header(title = "Optimal Cut-off Point Metrics - Diagnostic Odds Ratio") %>%
   fmt_number(
     columns = where(is.numeric),
     decimals = 3
   )

```

&nbsp; 
&nbsp; 

### Min p-value
```{r echo=FALSE, message=FALSE, warning=FALSE}
minpvalue_res <- optimal.cutpoints(
  X = "y_pred", 
  status = "x",
  tag.healthy = 0, 
  methods = "MinPvalue", 
  data = d_cut,
  ci.fit = TRUE, 
  conf.level = 0.95, 
  trace = FALSE)


summary(minpvalue_res)

# Results
opt_minp <- summary(minpvalue_res)$MinPvalue$Global$optimal.cutoff

cutoff_minp <- opt_minp$cutoff

se_minp<-opt_minp$Se
sp_minp<-opt_minp$Sp
ppv_minp<-opt_minp$PPV
npv_minp<-opt_minp$NPV


# Càlculs CI
sensitivity <- se_minp[1, "Value"]
sensitivity_low <-se_minp[1, "ll"]
sensitivity_high <- se_minp[1, "ul"]

specificity <- sp_minp[1, "Value"]
specificity_low <- sp_minp[1, "ll"]
specificity_high <- sp_minp[1, "ul"]
  
ppv <- ppv_minp[1, "Value"]
ppv_low <- ppv_minp[1, "ll"]
ppv_high <- ppv_minp[1, "ul"]
  
npv <- npv_minp[1, "Value"]
npv_low <- npv_minp[1, "ll"]
npv_high <- npv_minp[1, "ul"]


# Classification

# Confusion matrix
d$y_cat <- ifelse(d$y_pred >= cutoff_minp, "Predicted disease", "Predicted non-disease")


tab <- table(d$y_cat, d$x_c)
tab <- tab[,c(2,1)]

kable(tab, caption = "Contingency Table - Min p-value", align = "c") %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                 full_width = F, position = "center")

# Diagnostic
diag <- epi.tests(tab)
diag


# Accuracy
correct<-sum(tab[1,1], tab[2,2])

total <- sum(tab)

acc <- as.numeric(correct/total)
ci_acc <- binom.test(correct, total, conf.level = 0.95)$conf.int



# Metrics
metrics_minp <- data.frame(
  method = "Min p-value",
  optimal_cutpoint = cutoff_minp,
  
  sensitivity = sensitivity,
  sensitivity_low = sensitivity_low,
  sensitivity_high = sensitivity_high,
  
  specificity = specificity,
  specificity_low = specificity_low,
  specificity_high = specificity_high,
  
  ppv = ppv,
  ppv_low = ppv_low,
  ppv_high = ppv_high,
  
  npv = npv,
  npv_low = npv_low,
  npv_high = npv_high,
  
  accuracy = acc,
  accuracy_low = ci_acc[1],
  accuracy_high = ci_acc[2]
)

metrics_t<-metrics_minp %>%
   select(method, optimal_cutpoint, sensitivity, sensitivity_low, sensitivity_high,
          specificity, specificity_low, specificity_high, ppv, ppv_low, ppv_high, npv, npv_low, npv_high, accuracy, accuracy_low, accuracy_high) %>%
   gt() %>%
   tab_header(title = "Optimal Cut-off Point Metrics - Min p-value") %>%
   fmt_number(
     columns = where(is.numeric),
     decimals = 3
   )

```
&nbsp; 
&nbsp; 


## Comparison table

`r tab_cap("Comparison of Cut-off Methods: Small scenario")`

```{r echo=FALSE, message=FALSE, warning=FALSE}

tab_cutoffs <- dplyr::bind_rows(
  metrics_boot_expect |> mutate(method = "New approach: Expected value"),
  metrics_boot_maxdens |> mutate(method = "New approach: Maximum density"),
  metrics_youden |> mutate(method = "Youden's J statistic"),
  metrics_eucl |> mutate(method = "Euclidean distance"),
  metrics_mprod |> mutate(method = "Max. product Se*Sp"),
  metrics_iu |> mutate(method = "Index of Union"),
  metrics_cost |> mutate(method = "Cost approach"),
  metrics_misscost |> mutate(method = "Missclassification cost"),
  metrics_dor |> mutate(method = "DOR"),
  metrics_minp |> mutate(method = "Min p-value")
) 


# Format CI
format_ci <- function(value, low, high) {
  if (is.na(low) | is.na(high)) {
    return(sprintf("%.2f", value))
  } else {
    return(sprintf("%.2f [%.2f; %.2f]", value, low, high))
  }
}

# Metrics [CI]
tab_cutoffs_formatted <- tab_cutoffs |>
  mutate(
    Cutoff = round(optimal_cutpoint, 2), 
    Sensitivity = mapply(format_ci, sensitivity, sensitivity_low, sensitivity_high),
    Specificity = mapply(format_ci, specificity, specificity_low, specificity_high),
    PPV = mapply(format_ci, ppv, ppv_low, ppv_high),
    NPV = mapply(format_ci, npv, npv_low, npv_high),
    Accuracy = mapply(format_ci, accuracy, accuracy_low, accuracy_high
  )) |>
  select(
    Method = method, 
    Cutoff,
    Sensitivity,
    Specificity,
    PPV,
    NPV,
    Accuracy
  )


# Final comparison table
library(gt)

tab_cutoffs_formatted |>
  gt()|>
  tab_header(
    title = md("**Comparison of Cut-off Methods**")
  ) |>
  cols_label(
    Method = "Method",
    Cutoff = "Cut-off point",
    Sensitivity = "Sensitivity",
    Specificity = "Specificity",
    PPV = "PPV",
    NPV = "NPV",
    Accuracy = "Accuracy"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),  
    locations = cells_column_labels(everything())
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),  
    locations = cells_body(columns = Method)
  )



# Guardar cutoffs
# method_df <- data.frame(Method = tab_cutoffs_formatted$Method)
# save(method_df, file = "method_df_small_v2.Rdata")

load("method_df_small_v2.Rdata")

df <- tab_cutoffs_formatted |> dplyr::distinct(Method, .keep_all = T) |> dplyr::select(Cutoff)


method_df <- method_df|> dplyr::bind_cols(df)
rownames(method_df)<-NULL
save(method_df, file = "method_df_small_v2.Rdata")
```


